--- 
title: 'Protocols'
output:
  bookdown::html_document2: 
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
link-citations: yes
fontsize: 12pt
bibliography: References.bib
---

```{r packages, echo=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(rmarkdown)
library(bookdown)
#library(distill)
library(knitcitations)
library(formatR)
library(devtools)
library(usethis)
library(lubridate)
library(scales)

#Generate BibTex citation file for all R packages used to produce report
knitr::write_bib(.packages(), file = 'packages.bib')
```


```{r setup, include=FALSE, cache=FALSE, message = FALSE}
#Load packages
library(knitr)
library(rmarkdown)
# Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
#opts_chunk$set(tidy = TRUE, R.options = list(width = 60), highlight = TRUE)

## Caching code
opts_chunk$set(cache = 2, cache.path = "cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
#opts_chunk$set(fig.pos = "H")
```


# Chromosome counting

## Objective

Counting chromosomes of *Artemisia tridentata* using root meristems of individual lines maintained *in vitro* at BSU. This approach will help ascertaining the ploidy level (2n) of our individual lines.

## Material & reagents

**Day 1**

- Little cooler for getting ice (room XXX).
- Labeled glass large Petri dishes.
- Tweezers (to pull roots out of Magenta vessels).
- Distilled water.
- Space in the fridge to store at 4<span>&deg;</span>C for at least 10 hours (can do 24 hours).


## Plant material

Young whitish roots are sampled from individuals cultured *in vitro* either at <span style="color: black;font-weight:bold">Growth</span> or <span style="color: blue;font-weight:bold">Rooting</span> phases. 

## Sampling and fixing tissue

1. Sample young whitish roots and place them in a Petri dish.
2. Use achromatic spindle inhibitors to stop mitotic divisions (at metaphase) and contract chromosomes. This can be done following two approaches: 
    - Treat roots with saturated solution of water and <span>&#x3B1;</span>-monobromonaphtalene for 3 hours at room temperature. Start by homogenizing the stock solution and adding supernatant to cover roots in the Petri dish. 
    - Treat roots in a mix of water and crushed ice (4<span>&deg;</span>C) for 10 hours.
3. Plant tissue fixation and conserving chromosomes. 
    - Prepare a falcon tube with the solution of ethanol:acetic acid (3:1). Add few drops of the acetocarmin solution and one drop of iron acetate. Homogenize the solution by gently shaking the falcon tube.
    - Transfer roots into falcon tube and store at room temperature for 1-2 days. 

## Staining tissue

4. Pour acetocarmin solution in a porcelain crucible and add roots. Cover crucible with a lid.
5. Prepare a Petri dish filled with 45% acetic acid. This will be used to cool down stained roots and preserve them until chromosome numbers is established.
6. Gently heat solution in porcelain crucible with a Bunsen burner for ca. 2 minutes. Avoid boiling! This will stain chromosomes in red.
7. Pour content of porcelain crucible into Petri dish.    

## Squashing tissue

8. Put a drop of the solution in the Petri dish on a microscope slide and add a root tip in it. 
9. Remove the root cap under the scope using tweezers.This will facilitate the squash and observation of chromosomes.
10. Cover with a plastic coverslip and use handle of tweezers to squash the tissue. Go gently! 
11. Observe under microscope and take picture to validate chromosome count.

## Preparing stock solutions

Recipes to prepare solutions presented here were adapted from this [document](Data/Caryologie_methodes_INRA_091118.pdf) (in french).

### Ethanol:acetic acid solution

- Mix one volume of acetic acid for three volumes of ethanol. Store at 4<span>&deg;</span>C.

### Acetocarmin solution

For 1000 mL for acetocarmin solution: 

1. Measure 450 mL of 45% glacial acetic acid and pour in **XXX** (a vessel that we can heat). See Figure in this [document](Data/Carmin_acétique_acétate_de_fer.pdf).
2. Measure 550 mL of distilled water and add to vessel.
3. Weigh 5 gr. of carmin (Fisher C579-25). Carmine is a basic dye that is prepared from the insect *Coccus cacti*.
4. Add boileezers, and reflux for 2-3 hours.
5. Filter into dark bottles and store at 4<span>&deg;</span>C.

This solution can be stored for a long time. Staining can be intensified by adding ferric chloride (FeCl~2~·6H~2~O); add 5 mL of a 10 % ferric chloride solution per 100 mL of % acetocarmine.

Kansas State University as also put together a slightly different protocol for the acetocarmin solution: https://www.k-state.edu/wgrc/electronic_lab/aceto_stain.html

# Flow cytometry

## Objective

Estimate 2C genome size of *Artemisia tridentata* based on leaf material using the one-step protocol with general purpose buffer (GPB) by @Loureiro2007 and the protocol described in @Pellicer2014. The publications are accessible [here](Data/). 

The flow cytometry will be conducted at BSU using a [BD Accuri™ C6 Flow Cytometer](Data/bd_accuri_optical_filter_guide.pdf).

## Challenge

One thing that we need to bear in mind is that the coefficient of variation of the fluorescence peaks should be below 5%, and that the measurements should be taken using a linear scale rather that a logarithmic scale.

## Plant material

Genome size is estimated using **young, fresh leaves for both our target (sagebrush) and internal standards**. 

Concerning reference internal standards, we recommend that several species, covering a broad range of genome sizes, are kept growing in the laboratory to enable the most appropriate standard to be selected for each particular analysis.

### Internal standards

Check what seeds we have in the lab.:

- *Petroselinum crispum* ‘Curled moss’ (2C=4.5 pg).
- *Pisum sativum* ‘Ctirad’ (2C=9.09 pg).
- *Solanum lycopersicum* L. ‘Stupiké polní rane’ (2C=2 pg).


## Protocol

The recipe of the buffer and the steps are described in the paper, but briefly:

- Work under cold conditions (keep samples and buffers on ice).
- Cut a small portion of leaf tissue (1cm^2^, or less) from both the target and the internal standard.
- Place samples in a small petri dish with 1ml of ice-cold buffer.
- Chop with a sharp new razor blade till obtaining a fine lysate.
- Add another 1ml of buffer.
- Filter through a 30 um nylon mesh.
- Stain with 100 ul of 1mg/ml Propidium iodide. Nuclei were stained with 50 µg mL−1 propidium iodide (PI; Fluka, Buchs, Switzerland), and 50 µg mL−1 RNase (Sigma, St Louis, MO, USA) was added to nuclear suspension to prevent staining of double-stranded RNA.
- Incubate 5-10 min on ice.
- Measure on the flow cytometer (here the [BD Accuri™ C6 Flow Cytometer](Data/bd_accuri_optical_filter_guide.pdf)).

## Preparing buffer

### General composition of the lysis buffer

**GBP:** 0.5 mM spermine.4HCl, 30 mM sodium citrate, 20 mM MOPS (MOPS, 4-Morpholinepropane sulfonate), 80 mM KCl, 20 mM NaCl, 0.5 % (v/v) Triton X-100, pH 7.0.

### Procedure to prepare the lysis buffer (GBP)

1. Weigh the following chemicals into a glass (final volume of 200 mL):
    - Spermine 0.5 mM (0.0348 g)
    - Sodium citrate 3mM (1.764 g)
    - MOPS 20 mM (0.836 g)
    - KCl 80 mM (1.193 g)
    - NaCl 20 mM (0.233 g)
2. Add double distilled water to 200 mL and mix with a rotating magnet and dissolve the chemicals
3. Add 1 mL of Triton X-100 and mix until diluted
4. Adjust to pH 7
5. Add 6 g of PVP-40 and mix until dissolved (Add PVP while the buffer is being mixed as it tends to clog. Be patient and leave it with a rotating magnet until completely dissolved)
6. Store in the fridge or prepare aliquots of 50 mL which can be stored on the freezer. Use one each time and once thawed do not freeze again. It can be stored in the fridge for weeks.

The buffer is now ready to use. If you notice that the sample becomes dark when prepared, add 200 ul of B-mercaptoethanol to the buffer and mix (or equivalent if aliquots have been prepared).

# Managing Linux accounts and sharing folders

To more efficiently benefit from our Linux computers, we will provide overview of protocols to:

- Create users.
- Delete users.
- Create groups.
- Add users to groups.
- Create shared folders.
- Assign shared folders to a group.
- Switch user to check changes.

In RStudio, you can send commands directly into the Terminal (instead of the Console) by using this combination of keys: `Ctrl + Alt + Enter`. The Terminal has to be selected before using this approach.

## Create users

For this purpose, we will be using the `adduser` command:

```{bash, eval=F}
#Create user, add password and create home directory (without admin rights)
sudo adduser username

#For example
sudo adduser anthonymelton
```

You will be asked to answer multiple questions, only the password is mandatory.

To view the list of users on the Linux computer type this command:

```{bash, eval=F}
cut -d : -f 1 /etc/passwd
```

To view the list of logged in users do:

```{bash, eval=F}
w
```

To **provide admin (`sudo`) rights to a user** that you created do the following:

```{bash, eval=F}
#Create a new used with admin/sudo rights
sudo usermod -aG sudo username

#For example
sudo usermod -aG sudo anthonymelton
```

On the other hand, **if you want to reverse a user from `sudo` to `standard` rights**, do the following:

```{bash, eval=F}
#General syntax
sudo deluser username sudo
```

This command will only remove user 'username' from the sudo group, but it will not delete the user permanently from the system.  Now, they become regular user and cannot do any administrative tasks as sudo user.

## Delete users

After someone leaves the lab, you might have to delete their user account. You can do that by using the following command, but please make sure that you have archived all their data and code prior to proceeding.

```{bash, eval=F}
#If you want to delete a user and its home directory and mail spool
sudo deluser --remove-home username
```

## Create groups

To create a group, here `genomics` type the following command:

```{bash, eval=F}
#Create a group entitled `genomics`
sudo groupadd genomics
```

## Add users to groups

To add users to the groups (here `genomics`) type the following command:

```{bash, eval=F}
#Add users to the group, here `genomics`
sudo usermod -a -G genomics userID

# Examples
sudo usermod -a -G genomics svenbuerki
sudo usermod -a -G genomics bio_11
sudo usermod -a -G genomics anthonymelton
```

To know which users are already members of a group type the following command:

```{bash, eval=F}
grep genomics /etc/group
```

## Create shared folders

When multiple users need accessing to the same data, it is best practice to create shared folders and assign them to a group (made of multiple users, here the `genomics` group). 

To create the folder to be shared (here `Genomics_shared`) type the following command:

```{bash, eval=F}
#General syntax
sudo mkdir -p /path/Shared_Folder_ID

#Example (saved in home/)
sudo mkdir -p /home/Genomics_shared
```

## Assign shared folders to a group

Here, we assign a group (here `genomics`) to become owner of the shared folder (here `/home/Genomics_shared`) and give them appropriate permissions, which are also inherited by the subdirectories. This is done as follows:

```{bash, eval=F}
sudo chgrp -R genomics /home/Genomics_shared
sudo chmod -R 2775 /home/Genomics_shared
```

## Switch user to check changes

To switch user on Ubuntu you need to type the following command:

```{bash, eval=F}
#General syntax
su USER_name

#Example (you will have to provide password for bio_11)
su bio_11
```

# References

