--- 
title: 'Genome sequencing, phasing and annotation'
output:
  bookdown::html_document2: 
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
link-citations: yes
fontsize: 12pt
bibliography: References.bib
---

```{r packages, echo=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(rmarkdown)
library(bookdown)
#library(distill)
library(knitcitations)
library(formatR)
library(devtools)
library(usethis)

#Generate BibTex citation file for all R packages used to produce report
knitr::write_bib(.packages(), file = 'packages.bib')
```


```{r setup, include=FALSE, cache=FALSE, message = FALSE}
#Load packages
library(knitr)
library(rmarkdown)
# Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
#opts_chunk$set(tidy = TRUE, R.options = list(width = 60), highlight = TRUE)

## Caching code
opts_chunk$set(cache = 2, cache.path = "cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
#opts_chunk$set(fig.pos = "H")
```

# Overview

Summaries of procedures applied in this project to sequence, phase and annotate the sagebrush genome based on the `G2_b24_1` individual line is presented here. Click [here](Individual_lines.html) to learn more about `G2_b24_1`. We are briefly presenting below our predicted sequence data and biomass requirements for each sequencing technology. 

## Sequence data

Our sequencing and assembly strategy is described in section \@ref(wetlab), but a summary of sequence data and their associated NGS platforms is provided below (see also Table \@ref(tab:tabproject)):

- [**Whole Genome Shotgun**](https://www.ncbi.nlm.nih.gov/genbank/wgs/): 
    - Illumina HiSeq platform: To infer genome size and complexity (and produce a haploid draft genome).
    - PacBio Sequel II (using 8M SMRT cells): To conduct *de novo* assembly.
- **Proximity ligation**: 
    - [Omni-C libraries](https://dovetailgenomics.com/intro-omni-c-tech/) (sequenced on and Illumina platform): To conduct phasing.
- **RNASeq**: 
    - Illumina NovaSeq platform: To perform genome annotation.
- [**SRA Experiments**](https://www.ncbi.nlm.nih.gov/sra/): Sequence Read Archive (SRA) data will be made available on NCBI upon submission of the genome.
- **Protein Sequences**: Protein sequences will be made available.
- **Assembly/Genome**: An assembly at pseudo-chromosome level will be made available on NCBI upon submission of the genome.

## Biomass requirements per sequencing technology

We are estimating that ca. 120 gr of leaf biomass are necessary for genome sequencing, phasing and annotation. Table \@ref(tab:tabproject) provides a summary of sequencing technologies applied in this project, their purpose and biomass requirements. These data do not account for a preliminary DNA extraction trial (to conduct PacBio sequencing). 

```{r tabproject, echo=FALSE, eval=TRUE}
#Table with data on sequencing and number of plantlets required for project
dat <- data.frame(Type = c("Illumina HiSeq","PacBio sequencing", "Proximity ligation (Hi-C for phasing genome)", "RNA-seq"),
                    #Add purpose
                    Purpose = c("Genome size and complexity (incl. haploid draft genome)","*De novo* genome assembly", "Phasing", "Annotation"),
                    # add requirements 
                    `Unit/Biomass` = c("1 Illumina library = 20 mg","1 cell = 20 gr", "1 Illumina library = 6 gr", "1 library = 20 mg"),
                    # add Total for project
                    `Total units / biomass` = c("1 library = 20 mg", "5 cells = 100 gr", "3 libraries = 18 gr", "1 library = 20 mg"),
                    #Add Number of plantlets
                    `Number plantlets` = c(ceiling(0.2/0.8),ceiling(100/0.8), ceiling(18/0.8), ceiling(0.2/0.8))
                  )
colnames(dat) <- c("Type","Purpose","Unit/Biomass","Total units / biomass","Number of plantlets")
#Plot table
knitr::kable(dat, caption = "Summary of sequencing technologies, their purpose and requirements to complete the sagebrush genome project.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

In the case of the Illumina HiSeq, although we will only produce 1 library, it will be dispatched on five HiSeq runs.  

## Sequencing data by technology

The amount of data (bp) produced per sequencing technology is provided in Table \@ref(tab:dataseq) together with estimates of haploid genome coverage (x). 

```{r dataseq, echo=FALSE, eval=TRUE}
#Table with data on sequencing and number of plantlets required for project
dat <- data.frame(Type = c("Illumina HiSeq","PacBio sequencing", "Proximity ligation (Hi-C for phasing genome)"),
                    #Add purpose
                    Purpose = c("Genome size and complexity (incl. haploid draft genome)","*De novo* genome assembly", "Phasing"),
                    # add requirements 
                    `Data/run` = c(350e6*300, 50e9, 100e6*300),
                    #N runs
                    `N runs` = c(5,5,3)
                  )
#Add column with total data generated
dat$Total <- dat$Data.run * dat$N.runs

#Add column with haploid genome coverage
dat$Coverage <- round(dat$Total/4.5e9,2)

colnames(dat) <- c("Type","Purpose","Data (bp)/Run","N. runs","Total data (bp)", "Haploid genome coverage (x)")
#Plot table
knitr::kable(dat, caption = "Summary of sequencing data produced by technology for the sagebrush genome project. Haploid genome size = 4.5Gbp") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


# Wet-lab procedures {#wetlab}

Completion of the wet-lab work by Dovetail Genomics detailed here is predicted to take 28 weeks (7 months) upon receipt of the biomass. To gain insights into the timetable for biomass production, please click [here](Biomass_production.html).

1. Ascertain the ploidy level, genome size and genome complexity of `G2_b24_1`. This will be done by counting chromosomes (based on root squashes), inferring 2C genome size using flow cytometry (based on root and leaf tissues) and estimating genome size and complexity by applying a k-mer approach on Illumina HiSeq data. 
2. Dovetail Genomics will extract HMW DNA with an average fragment size ca. 50-100kb based on biomass supplied by the team (see [Biomass production](Biomass_production.html)).
3. Whole genome sequencing (WGS) will be conducted using PacBio technology. We aim at generating 50-100X raw data coverage (haploid sagebrush genome size: 4.5 Gbp). To achieve this objective, 5 PacBio CLR (continuous long-read) libraries will be built and sequenced on the PacBio Sequel II using 8M SMRT cells. A SMRT cell should generate >50Gb of raw data. In our case, 250Gb (5x 50Gb) of raw data will be generated, which corresponds to 55X haploid genome coverage (we are on the low side, but it should be fine). These data will be used to assemble the haploid draft genome (used as input for phasing based on Omni-C libraries).
4. Three [Omni-C libraries](https://dovetailgenomics.com/intro-omni-c-tech/) (= proximity ligation libraries) will be constructed (one library per 3 Gbp of the organism’s genome) based on the HMW DNA extractions. These libraries will be sequenced on an Illumina HiSeq X instrument (ca. 100M PE150bp read pairs per Gbp of genome size). Before final sequencing, Dovetail will conduct some shallow sequencing (ca. 2M PE75 bp) to assess library quality.
5. RNA will be extracted from leaf tissue (20mg) by Dovetail and a standard RNA library will be prepared with rRNA-depletion. 2x150bp read will be sequenced on an Illumina platform and raw data from ca. 20M read pairs per sample (1 in this project) will be used for genome annotation.  

# Bioinformatic procedures

1. PacBio *de novo* assembly (= produce haploid draft assembly). We aim at a minimum N50 of 100kb (needed for HiRise pipeline).
2. The haploid daft assembly will be scaffolded (phased) through the HiRise software pipeline using the proximity ligation data (this software is owned by Dovetail). The Omni-C libraries can be used for genome assembly and haplotype phasing. We will also be able to use this data to call SNPs and look at structural variations.
3. Whole genome annotation will be conducted using RNA-seq data. The basic services will be included:
    a. Repeat masking.
    b. *Ab initio* gene prediction using related species (e.g. *Artemisia annua*).
    c. RNAseq mapping to enhance annotation.
    d. Manual curation of 5 genes (not very interesting for us). The work done by Anthony Melton based on seedlings from GxE experiment will be key here.
    e. Assignment of functional tags to genes.
4. Results files will be delivered electronically via secure FTP.


<div id="refs"></div>

# Appendix 1 {.appendix}

Citations of all R packages used to generate this report. 

```{r generateBibliography, eval = T, results="asis", cache = F, echo=F, warning = FALSE, message=FALSE}
library("knitcitations")
cleanbib()
options("citation_format" = "pandoc")
read.bibtex(file = "packages.bib")
``` 

# Appendix 2 {.appendix}

Version information about R, the operating system (OS) and attached or R loaded packages. This appendix was generated using `sessionInfo()`.

```{r eval=T, echo=F, warning = FALSE, message=FALSE}
sessionInfo()
```
